---
title: "Visualising and Modelling Data"
author: "Simon Parker"
date: "21 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A simple chart 

```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point()
```

Now we will add a bit of colour

```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point(aes(color = manufacturer))

```

I'm interested to see whether there is a difference between the fuel consumption of cars with large or small engines. To test this I will use dplyr to mutate a new variable from the mpg data, based on the displacement variable.

```{r}
mpg2 = mpg %>% 
  mutate(size = cut(displ, 3, labels = c("Small", "Medium", "Large"))) 
  
ggplot(mpg2, aes(cty, hwy)) + geom_point(aes(color = size))

```

We can add further information to our chart by simply changing the aesthetic mapping.

```{r}
mpg2 = mpg %>% 
  mutate(size = cut(displ, 3, labels = c("Small", "Medium", "Large"))) 
  
ggplot(mpg2, aes(cty, hwy)) + geom_point(aes(color = manufacturer, size = size))

```

I think we have made this a little tricky to read, it may be worth removing the size attribute from the mapping. Instead we can use facet to divide our graph into groups.

```{r}
mpg2 = mpg %>% 
  mutate(size = cut(displ, 3, labels = c("Small", "Medium", "Large"))) 
  
ggplot(mpg2, aes(cty, hwy)) + geom_point(aes(color = manufacturer)) + facet_wrap("size")

```

We can then try change visualisation to layer, which may radically alter our graph.

For example: 

```{r}
mpg2 = mpg %>% 
  mutate(size = cut(displ, 3, labels = c("Small", "Medium", "Large"))) 
  
ggplot(mpg2, aes(cty, hwy)) + geom_smooth() + facet_wrap("size")

```

Or maybe a boxplot would be useful.

```{r}
mpg2 = mpg %>% 
  mutate(size = cut(displ, 3, labels = c("Small", "Medium", "Large"))) 
  
ggplot(mpg2, aes(size, cty)) + geom_boxplot(fill = "lightcyan1") + theme(legend.position="none")


```

I want to find out if there is a difference in the highway (motorway) fuel economy of these cars when looking at the different transmissions, and how this has changed over time. The issue with simply using a facet of the trans variable, is that there are multiple auto and manual transmissions. 


Using grepl and case_when, we can group up strings into new categories. In the example below I first mutate a new AM variable, then group up the different types of auto transmissions into a category named 'Auto'. As there are multiple types of manual transmission I also create a 'Manual' group.

```{r}
mpg3 <- mpg %>%  
  mutate(AM = case_when(grepl("auto", trans) ~ "Auto", grepl("manual", trans) ~"Manual"))
  
  ggplot(mpg3, aes(AM, hwy)) + geom_boxplot(aes(fill = AM)) + theme(legend.position="none") + facet_wrap("year")
```

# Exercise 1

```{r}

```

Visualising Sparklyr

```{r}
library(sparklyr)
library(dplyr)
library(nycflights13)
library(ggplot2)

sc <- spark_connect(master="local")
flights <- copy_to(sc, flights, "flights", overwrite = T)
airlines <- copy_to(sc, airlines, "airlines", overwrite = T)
planes <- copy_to(sc, planes, "planes", overwrite = T)
weather <- copy_to(sc, weather, "weather", overwrite= T)

src_tbls(sc)

```

All we need to do is use the collect() command.

```{r}

flights %>%
select(year, month, day, dep_delay, carrier) %>%
filter(month == 12, day == 25) %>%
select(carrier, dep_delay) %>%
arrange(desc(dep_delay)) %>%
collect() %>%
ggplot(aes(dep_delay)) + geom_bar(aes(fill= carrier))  
```
  
# Exercise 2

Without loading the data into R, present a chart that you think is interesting based on the nycflights13 data

```{r}

```

